Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-03-17 paper.tex
section 7

These coincidences should not be a surprise since existing literature is extraordinarily rich
in showing the correspondence between monads and continuation-passing style as well as
delimited controls since their very beginning era. \todo{cite}

@@

\subsection{Un-CPS by Delimited Controls} \label{uncps}

As identified before, \note{JD: compelling...}

\begin{lstlisting}
\end{lstlisting}

\iffalse
\section{Refunctionalized AAM}
Maybe insert to section refunc.
\subsection{Store Widening}
\subsection{Precise Call/Return Match}
show the precision of stack
On equivalence: since refunc interp does not have an explicit concept of $state$,
so 1) we can compare the final states of P4F against with refunc interp,
or 2) we can compare the summarized store of P4F against with summarized store of refunc interp.
\subsection{Termination}
show that number of continuation is finite.
\subsection{Soundness}
\subsection{Complexity}
\fi

@@

\iffalse
The following code shows the implementation:
\begin{lstlisting}
case class P4FContAddr(tgt: Expr, tgtEnv: Env) extends KAddr
def allocKont(tgtExpr: Expr, tgtEnv: Env): KAddr = P4FContAddr(tgtExpr, tgtEnv)
\end{lstlisting}
The target environment is used for successor states and extended from the closure's
environment (namely @val new_env = env_c + (v -> baddr)@).
\fi

>>>>

These coincidences should not be a surprise since existing literature is extraordinarily rich
in showing the correspondence between monads and continuation-passing style as well as
delimited controls since their very beginning era \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.

@@

\todo{Update the code}
\subsection{Un-CPS by Delimited Controls} \label{uncps}

As identified before, there are two layers of continuations, one for the normal stack of 
the analyzed language, and the other for the nondeterministic choices of closures.
This corresponds to the extended continuation-passing style.
But in our defining language Scala, the delimited controls rely on the continuation-passing
transformation which is only one layer.
To address this problem, base on the @nd@ operator we have before, we introduce 
another operator @ndcps@ for handling non-deterministic choices.

The @ndcps@ itself is implemented with delimited controls. 
The function takes two arguments: a set of elements with type @T@ and an initial 
accumulated value of type @Set[S]@. 
Function @ndcps@ iteratively returns a value of type @T@ with annotated by
@@cps[Set[T]]@ that denotes the final returned value of @ndcps@ is type @Set[T]@.
The @f@ introduced by the @shift@ operator is the delimited continuation at the 
call-site of @ndcps@.

\begin{lstlisting}
def nd[T,S](ts: Set[T], acc: Set[S], k: T => Set[S]): Set[S] = {
  if (ts.isEmpty) acc
  else {
    val vss = k(ts.head)
    nd(ts.tail, acc ++ vss, k)
  }
}

def ndcps[T,S](ts: Set[T], acc: Set[S]): T @cps[Set[S]] = shift { f: (T => Set[S]) =>
  nd(ts, acc, f)
}
\end{lstlisting}

We add an annotation @@cps[Ans]@ on the return type of function @aeval@.
Now for every non-deterministic choices in the abstract interpreter

\begin{lstlisting}
def aeval(e: Expr, env: Env, store: BStore, time: Time): Ans @cps[Ans] = {
  val new_time = (e::time).take(k)
  e match {
    case Let(x, App(f, ae), e) =>
      val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
      val Clos(Lam(v, body), c_env) = ndcps[Clos, VS](closures, Set[VS]())
      val vbaddr = allocBind(v, new_time)
      val new_cenv = c_env + (v -> vbaddr)
      val new_store = store.update(vbaddr, aeval(ae, env, store))
      val bodyvss = aeval(body, new_cenv, new_store, new_time)
      val VS(vals, time, store) = ndcps[VS, VS](bodyvss, Set[VS]())
      val xbaddr = allocBind(x, time)
      val new_env = env + (x -> xbaddr)
      val new_vsstore = store.update(xbaddr, vals)
      aeval(e, new_env, new_vsstore, time)

    case ae if isAtomic(ae) =>
      val ds = atomicEval(ae, env, store)
      Set(VS(ds, new_time, store))
  }
}
\end{lstlisting}

@@
